Baseline centering problem

### 2018-12-21 paper/tangotd.tex

The above can be turned to an infinite-rank approximation by writing
\begin{equation}
\transp{1_{s_1}}Z\psi=\E_{s_2\sim \rho}\, \E_{\omega}\,
\green(s_1,\omega)f(s_2,\omega)\psi(s_2)
\end{equation}
where integration is over $\omega$ in an arbitrary space (eg $\omega\in
[0;1]$ or $\omega\in \R^d$ with Gaussian distribution). This amounts to
sending each state to a function over $\omega$, and using the dot product
from $L^2(\omega)$. This should solve some kernel/image issues and
increase representational power. The equation
$\Id=F\Delta\transp{\green}$ becomes $\delta_{\omega\omega'}=\E_{s\sim
\rho, \,s'\sim P(s'|s)}
f(s,\omega)(\green(s,\omega')-\green(s',\omega'))$ TODO CHECK, which is not easy to
solve because of the singular $\delta_{\omega\omega'}$, but should be better with things on the right and left.

@@

>>>>

The above can be turned to an infinite-rank approximation by writing
\begin{equation}
\transp{1_{s_1}}Z\psi=\E_{s_2\sim \rho}\, \E_{\omega}\,
\green(s_1,\omega)f(s_2,\omega)\psi(s_2)
\end{equation}
where integration is over $\omega$ in an arbitrary space (eg $\omega\in
[0;1]$ or $\omega\in \R^d$ with Gaussian distribution). This amounts to
sending each state to a function over $\omega$, and using the dot product
from $L^2(\omega)$. This should solve some kernel/image issues and
increase representational power. The equation
$\Id=F\Delta\transp{\green}$ becomes $\delta_{\omega\omega'}=\E_{s\sim
\rho, \,s'\sim P(s'|s)}
f(s,\omega)(\green(s,\omega')-\green(s',\omega'))$ TODO CHECK, which is not easy to
solve because of the singular $\delta_{\omega\omega'}$, but should be
better with things on the right and left. For instance, the Newton equation on
$\transp{\green}$ is
\begin{equation}
\green(s_1,\omega)\gets
(1+\eta)\green(s_1,\omega)-\eta\E_{s\to s'}\E_{\omega'}
\green(s_1,\omega')f(s,\omega')(\green(s,\omega)-\gamma
\green(s',\omega))
\end{equation}

@@

The Newton update $Z\gets 2Z-Z\Delta Z$ is stable provided it is
initialized to some $Z$ such that $\Id-\Delta Z$
has operator norm $< 1$. With $\Delta=\Id-\gamma P$, the initial value
$Z=\Id$ satisfies this provided $\gamma<1$ (and probably $\gamma=1$ for
the relative case). Here we use a stochastic update, which should follow
the deterministic dynamics if the learning rate is small enough. However,
in general we will not apply the update $Z\gets (1+\eta)Z-\eta Z\Delta
Z$, even in expectation, because we will also sample $s_1$ and $s_2$.
Sampling $s_1$ according to $\rho$ amounts to multiplying the update by
$\rho$. This leaves the fixed point unchanged, but may change the
stability TODO CHECK. Using a gradient descent on parameters will further
introduce a preconditioner on the update. TODO CHECK.

For discrete time but continuous state, the true $Z=\sum \gamma^t P^t$ is
equal to $\Id$ plus a continuous component; the $\Id$ is singular, so
learning $Z(s_1,s_2)$ using eg a neural network will be difficult around
the diagonal (and the initialization will be difficult). We may directly
write the dynamics on $Z-\Id$. (TODO not clear what this means in
continuous time: not clear if $\Delta^{-1}-\Id$ is actually small; still,
starting at $\Id$ is stable for Newton whatever $P$ is, so it should be
stable in the limit.)

Setting $Z=\Id+\zeta$ provides the update
\begin{equation}
\zeta\gets (1-\eta)\zeta+\eta\gamma P+\eta\gamma P\zeta
+\eta\gamma \zeta P-\eta\zeta(\Id-\gamma P)\zeta
\end{equation}
which we cannot quite sample (because of the $P$ term). But it mixes the
left and right Bellman equations nicely. A way to sample it is to sample
the left state $s_1$ according to $\rho$; this amounts to multiplying the
update by $\rho$ (same fixed point, but maybe different dynamics TODO),
and then $\rho P$ is samplable.

The convergence of the update above to its fixed point $\gamma
P(\Id-\gamma P)^{-1}$ can be studied by studying the
remainder $r\deq \zeta-\gamma
P(\Id-\gamma P)^{-1}$, whose dynamics is
\begin{equation}
r\gets (1-\eta) r-\eta r(\Id-\gamma P)r
\end{equation}
In particular, at first order for small $r$ this is just $r\gets
(1-\eta)r$. This is contracting \emph{independently of the properties of
$P$}, contrary to the Bellman equation. The stochastic
Newton update with learning rate $\eta$ does not have quadratic
convergence like the standard Newton method, but it still keeps better
local convergence properties than Bellman. Is it worth the cost?
